---
title: "Women in Work Force"
output: html_notebook
---

```{r} 
#Importing Data Set
ss14pusa = read.csv(file.choose(),header=TRUE)  # read csv file
ss14pusb = read.csv(file.choose(),header=TRUE)

#Libraries
require(plyr)
require(dplyr)
```

```{r}
#Importing Data Set in a faster way
library(data.table)
variables=c("SEX","OCCP","WKHP","MAR","WAGP","SCHL","ESP","RAC1P","ST","PAOC","PWGTP")
ss14pusa=fread("~/Google Drive/Data for R/csv_pus/ss14pusa.csv",head=TRUE,select=variables)
ss14pusb=fread("~/Google Drive/Data for R/csv_pus/ss14pusb.csv",head=TRUE,select=variables)

```

```{r}
attach(ss14pusa)
names(ss14pusa)
ss14pusa_edit = data.frame(SEX,OCCP,WKHP,MAR,WAGP,SCHL,ESP, RAC1P, ST, PAOC, PWGTP)
detach(ss14pusa)

attach(ss14pusb)
ss14pusb_edit = data.frame(SEX,OCCP,WKHP,MAR,WAGP,SCHL,ESP, RAC1P, ST, PAOC, PWGTP)
detach(ss14pusb)

Data = rbind(ss14pusa_edit,ss14pusb_edit)
colnames(Data) <- c("Gender","Occupation", "Work_hours", "Marriage", "Income", "Education" ,"Parental_Occupation", "Race", "State", "Children", "Weight")


write.csv(Data, file = "Data.csv",row.names=TRUE)
```

```{r}
#Delete income=0/NA rows
row_to_keep=which(Data$Income>0)
Data=Data[row_to_keep,]
```

```{r, echo = FALSE}
#recode OCCP
class(Data$Occupation) <- "numeric"
Data$Occupation <- ifelse(Data$Occupation >= 10 & Data$Occupation <= 430, 1, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 1005 & Data$Occupation <= 1240, 2, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 800 & Data$Occupation <= 950, 3, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 2105 & Data$Occupation <= 2160, 4, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 3000 & Data$Occupation <= 3540, 5, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 510 & Data$Occupation <= 740, 6, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 1300 & Data$Occupation <= 1560, 7, Data$Occupation)
Data$Occupation <- ifelse(Data$Occupation >= 1600 & Data$Occupation <= 1965, 8, Data$Occupation)

Data <- filter(Data, Occupation %in% c(1:8))
Data$Occupation <- as.factor(Data$Occupation)

levels(Data$Occupation) <- c('MGR', 'CMM', 'FIN', 'LGL', 'MED' , 'BUS', 'ENG', 'SCI')
```

```{r}
#Recode Region
#https://en.wikipedia.org/wiki/List_of_regions_of_the_United_States#/media/File:Census_Regions_and_Division_of_the_United_States.svg
require(car)
Data$Region <- recode(Data$State,"c(09,23,25,33,44,59,34,36,42)='Northeast';c(17,18,26,38,55,19,20,27,29,31,38,46,39)='Midwest'; c(10,12,13,24,37,45,51,11,54,01,21,28,47,05,22,40,48)='South';c(04,08,16,30,32,35,49,50,02,06,15,41,53,56)='West';'72'='Puerto Rico'")
```
```{r}
# Recode Division
Data$Division <- recode(Data$State,"c(53,41,6,2,15)='Pacific';c(4,8,16,30,32,35,49,56)='Mountain'; c(19,20,27,29,31,38,46)='West North Central';c(17,18,26,39,55)='East North Central';c(5,22,40,48)='West South Central';c(1,21,28,47)='East South Central';c(34,36,42)='Middle Adlantic';c(10,11,12,13,24,37,45,51,54)='South Atlantic';c(9,23,25,33,44,50)='New England'")
```

```{r}
#Recode Education
Data$Education <- recode(Data$Education,"c(01,02,03,04,05,06,07,08,09,10,11)='~greade8';c(12,13,14,15,16,17,18,19)='grade9~college_nodegree';c(20,21)='associate/bachelor';c(22,23)='master/professional';c(24)='doctor')

```

```{r}
#Recode Education
library(car)
Data$Education <- recode(Data$Education,"1:11='~greade8';12:19='grade9~college_nodegree';20:21='associate/bachelor';22:23='master/professional';24='doctor'")
```


```{r, echo = FALSE}
#add in weights...example code from previous project:
require(survey)

df1<-svrepdesign(variables=pus_new[,1:13], 
repweights=pus_new[,14:93], type="BRR",combined.weights=TRUE,
weights=pus_new$PWGTP)
summary(df1)
svymean(~ WAGP,df1, na.rm = T)

#If We only want to know the SE and Mean of Statistics Income?

pus_new$FOD1P[pus_new$FOD1P == 3702 | pus_new$FOD1P == 6212 | pus_new$FOD1P == 6202]<-"Statistics"
df<-subset(pus_new, pus_new$FOD1P=="Statistics")
df2<-svrepdesign(variables=df[,1:13], 
repweights=df[,14:93], type="BRR",combined.weights=TRUE,
weights=df$PWGTP)
summary(df2)
svymean(~ WAGP,df2, na.rm = T)
```

```{r, echo = FALSE}

#median wage for each occupation by gender (just to acknowledge that there *is* a gender gap)
require(ggplot2)
Data_sex_occp <- ddply(Data, .(Gender, Occupation), summarise, MEAN = weighted.mean(Income, Weight, na.rm = T))
ggplot(Data_sex_occp, aes(x=Occupation, y=MEAN, fill=factor(Gender))) + 
  geom_bar(stat="identity",position="dodge") + 
  scale_fill_brewer(palette="RdYlGn") +
  labs(fill="") + 
  ylab("Mean Salary ($)") + 
  xlab("Occupations") + 
  ggtitle(paste("Salary Comparison between Men & Women")) + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1), 
        panel.background = element_rect(fill = 'white' )) + 
  theme_grey(base_size = 12)

```

```{r, echo = FALSE}
#subset on women, explore groups of women
Data_women <- filter(Data, Gender == 2)
#Race - Catherine
#Motherhood - 
#Marital Status - Erica (potentially add in region)
#Proportion of graduate degrees in each occupation?
#Region - Vivien

```

```{r, echo = FALSE}
Data_women <- filter(Data, Gender == 2)
MotherHood=select(Data_women, Income,Children)
#detach(package:plyr)
GroupedMotherHood <-
  MotherHood %>%
  na.omit() %>%
  group_by(Children) %>% 
  summarize(
    AvgIncome = mean(Income)
  ) 


### Bar Chart
ggplot(GroupedMotherHood, aes(x=Children, y=AvgIncome)) + 
  geom_bar(colour="black", fill="#DD8888", width=.5, stat="identity") + 
  guides(fill=FALSE) +
  xlab("Children") + ylab("Mean Income") +
  coord_cartesian(ylim = c(50000,70000)) +
  #scale_y_continuous(limits = c(40000,70000),oob = rescale_none) +
  ggtitle("Income for Motherhood")



### 
```

```{r}
MotherWorking=select(Data_women, Income,Children,Work_hours)
#detach(package:plyr)
GroupedMotherWorking <-
  MotherWorking %>%
  na.omit() %>%
  group_by(Children,Work_hours) %>% 
  summarize(
    AvgIncome = mean(Income),
    count=n()
  ) 

library(plotly)
# note how size is automatically scaled and added as hover text

plot_ly(GroupedMotherWorking, x = Work_hours, y = AvgIncome,size=sqrt(count), color = Children,text = paste("Count: ", count),opacity=1-Children*0.2,mode = "markers")

#plot_ly(GroupedMotherWorking, x = Work_hours, y = AvgIncome, text = paste("Count: ", #count),mode = "markers", opacity=1-Children*0.2,color = Children)
```
Observations:
1. When work_hours smaller than 60Hrs, Avg Income tend to be positively related to work hour.
2. There are generally more females with no Children
3. Females with Children from 6 to 17 often has higher avg income
```{r}
MotherOccupation=select(Data_women, Income,Children,Occupation)
#detach(package:plyr)
GroupedMotherOccupation <-
  MotherOccupation %>%
  na.omit() %>%
  group_by(Children,Occupation) %>% 
  summarize(
    AvgIncome = mean(Income),
    count=n()
  ) 


revalue(GroupedMotherOccupation$Children,c('1'='Children under 6','2'='Children from 6 to 17','3'='Children under 6 and above 6','4'='No Children'))

ggplot(GroupedMotherOccupation,aes(x=Children))+geom_point(aes(y=Occupation,size=count,colour=AvgIncome))+scale_colour_gradient(low="green",high="red")+scale_x_continuous(breaks=1:4, labels=c("Children under 6", "Children from 6 to 17", "Children under 6 and above 6", "No Children")) 

#scale_colour_gradientn(colours = terrain.colors(10))
```
Observations:
1. Females with Children from 6 to 17 usually tend to have higher average income as the dot has darker color.
2. Occupation of ENG, MGR and CMM usually tend to have higher salary. On the other hand, female in LGL field always have lower salary
3. 

```{r}
# Regional Effect
# boxplot of each state/division/region to see distribution
# calculate median of women in each Division and run significance test?
# calculate median of women in each State and plot in mapview
DW=Data[Data$Gender==2,]

boxplot(DW$Income~DW$Region) 
plot(DW$Income~DW$Division)
DR=cbind(DW$Income,DW$Division)
head(DR)
hist(DR[DR[,2]=="Pacific"][,1])

library(dplyr)
DivMedian <- aggregate(Income~Division,Data_women, median)
RegMedian <- aggregate(Income~Region,Data_women, median)
DivMedian
RegMedian


xtabs(~Gender+Division, data=Data)
```



